---
title: "Assignment10"
author: "Erin Keller"
date: "May 2, 2017"
output: html_document
---
---
title: "Assignment10"
author: "Erin Keller"
date: "April 5, 2017"
output: html_document
---
# Question 1: Using for loops to determine # zeros in a numerical vector
```{r}
Vector <- c(0,0,1,1,0,1,0,1,1,1)
head(Vector)
Vec0<-vector(mode="logical")
for (i in 1:length(Vector)) {
  if (Vector[i] == 0) {
    Vec0[i] <- 1 
     }
  else {
    Vec0[i]<- 0
  }
}
  sum(Vec0)

```

# Question 2: Using subsetting rather than for loops
```{r}
Vector <- c(0,0,1,1,0,1,0,1,1,1)

length((Vec0) <= subset(Vector, Vector == 0))
```
# Question 3: Maximum difference between all pairs
### Script courtesy of Alex Looi, annnotations done by Erin L. Keller
```{r}
X = rnorm(20, mean = 0, sd = 10) # sampling 20 random numbers with mean of 0

max_diff = function(X){
  
  pairs = expand.grid(X,X) # identifying all pairwise comparisons in X
  diffs = abs(pairs$Var1 - pairs$Var2) # calculating the absolute difference between all pairs
  
  max_difference = max(diffs, na.rm = T) # determining max difference by using the max function to determine the largest diference present in the diffs vector
  
  return(max_difference) # returns the max difference
  
}
max_diff(X)

```

# Question 4: Modifying output of Question 3 to contain 3 elements: 1st is the pair of values that give the greatest difference, 2nd is the position of these elements, and 3rd is the maximum distance from this pair
```{r}
X = rnorm(20, mean = 0, sd = 10)

max_diff = function(X= rnorm(20, mean = 0, sd = 10)){
  
  pairs = as.matrix(expand.grid(X,X))
  diffs = abs(pairs[,1] - pairs[,2]) # up until this point in the script is the same as Q3
  
  vec_max = which.max(diffs) # this is selecting the maximum difference from all pairs
  
  pair_max = pairs[vec_max,] # this is identify which pair of values generated largest difference
  
  max_difference = max(diffs, na.rm = T) # This is pulling out the maximum difference between two values
  
  var1_vec = which(pair_max[1] == X) # this is determining the position of the first value of the pair
  var2_vec = which(pair_max[2] == X) # this is determinig the position of the second value of the pair
  pair_vec = c(var1_vec, var2_vec) # this is concatenating the above output and contains the position of these values
  
  output = list(pair_max, pair_vec, max_difference) # ouput will be a list containing the 3 (really 4) values
  
  return(output)
  
}
max_diff()
```
# Question 5: Keep only maximum difference but cycle through all possibilities
```{r}
X = rnorm(20, mean = 0, sd = 10)

max_diff = function(X= rnorm(20, mean = 0, sd = 10)){
  
  pairs = expand.grid(X,X) # making pairwise comparisons
  diffs = abs(pairs$Var1 - pairs$Var2) # calculating absolute difference between every pairwise comparison
  temp_max = 0 # creating a temporary maximum vector containing what will be the largest difference
  for(v in 1:length(diffs)){ # for loop in which each pairwise comparison is being evaluated (length(diffs) means that every difference value in the vector will be evaluated)
    
    if(temp_max <= diffs[v]){ # if statement saying that if the difference value being evaluated is less than or equal to what is already in the temp_max file
      temp_max = diffs[v] # so if this is true (new diff is the greater than what was previously stored) then we replace what is in temp_max with the new maximum difference
    }
  }
  return(temp_max)
}
max_diff(X)
```

# Question 6: Multiplying two matrices together
```{r}
M1 = matrix(c(1:6),2,3, byrow = T) # generating two matrices
M2 = matrix(c(7:12),3,2, byrow = T)

Mat_mult = function(M1, M2){ # creatig function that will take the two matrices as inputs and multiply them together
  # create a new matrix table
  M3 = matrix(0, nrow(M1), ncol(M2)) # making empty matrix with rows = # rows in M1 and columns = # columns in M2
  if (nrow(M1) != ncol(M2)){ # if the row and column lengths are not equal, give the following warning:
    cat("input matrices are not correct dimensions", "\n")
  } else { # if they are the same length, do the following:
    for (nr in 1:nrow(M1)){ # for the number of rows in 1:length of M1 and for the number of columns in 1:length of M2, take the product of the values at the corresponding position in the matrix
      for (nc in 1:ncol(M2)){
        cell_val = sum(M1[nr,]*M2[,nc])
        M3[nr,nc] = cell_val # output will be a matrix containing the product of the two matrices
      }
    }
  }
  return(M3)
}

Mat_mult(M1, M2)

# Using R funtion formultiplying matrices:
M1%*%M2 # returns same result!
```
# Question 7: 
